---
title: "Differential coexpression of multiple brain regions - Mt.Sinai array dataset"
author: "Sandeep Amberkar"
date: "03/01/2018"
output: 
    prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,eval = T)
```

## Analysis Summary

Differential coexpression (DCx) is the change in the coexpression levels of a gene pair between 2 conditions, namely, Case and Control. This principle has been applied in many diseases to identify transcriptional changes beyond the scope of differential expression. Using 'diffcoexp'(https://github.com/hidelab/diffcoexp), DCx analysis was performed on the Mt.Sinai array dataset comprising of 951 samples across 17 brain regions profiled on the Affymetrix hgu133 platform. 

### Results


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
